---
title: "Xiaoting(Theresa) Liu - Final Project"
output:
  html_document:
    df_print: paged
---

#Improting libraries:
```{r}
library(plyr)
library(MASS)
# data re-shaping
library(tidyr) 
# data re-shaping
library(dplyr)  
```

```{r}
library(ggplot2)    
library(knitr)
library(broom)
library(reshape2)
```

```{r}
# pipe operator
library(magrittr)   
```

#Descritive Analysis 
```{r}
#Import dataset
heartdiseases <- read.csv ("~/Desktop/Course Data/heart.csv",header=T)
head(heartdiseases)
```

```{r}
#Clean the dataset and summarize the dataset 
heartdiseases <- heartdiseases %>% drop_na() #drops rows with null values 
summary(is.na(heartdiseases))
```
The dataset is pretty well organized. There is no null data. 


```{r}
#View the variables and the type of the data
str(heartdiseases)
```

It is better to change the column names and value of the categorical variable for better interpretation later. 
```{r}
#change the column names
name <- c('age', 'sex','chest_pain_type', 'resting_blood_pressure', 'cholesterol', 'fasting_blood_sugar', 'rest_ecg', 'max_heart_rate_achieved','exercise_induced_angina', 'st_depression', 'st_slope', 'num_major_vessels', 'thalassemia', 'target')
names(heartdiseases) <- name
```


```{r}
#changes the value
heartdiseases$sex[heartdiseases$sex == "1"] = "male"
heartdiseases$sex[heartdiseases$sex == "0"] = "female"
heartdiseases$chest_pain_type[heartdiseases$chest_pain_type == "0"] = "typical angina"
heartdiseases$chest_pain_type[heartdiseases$chest_pain_type == "1"] = "atypical angina"
heartdiseases$chest_pain_type[heartdiseases$chest_pain_type == "2"] = "non-anginal pain"
heartdiseases$chest_pain_type[heartdiseases$chest_pain_type == "3"] = "asymptomatic"
heartdiseases$fasting_blood_sugar[heartdiseases$fasting_blood_sugar == "0"] = "<=120"
heartdiseases$fasting_blood_sugar[heartdiseases$fasting_blood_sugar == "1"] = ">120"
heartdiseases$rest_ecg[heartdiseases$rest_ecg == "0" ] = "Normal"
heartdiseases$rest_ecg[heartdiseases$rest_ecg == "1" ] = "Abnormality"
heartdiseases$rest_ecg[heartdiseases$rest_ecg == "2" ] = "Probable or Definite"
heartdiseases$exercise_induced_angina[heartdiseases$exercise_induced_angina == "0"] = "no"
heartdiseases$exercise_induced_angina[heartdiseases$exercise_induced_angina == "1"] = "yes"
heartdiseases$st_slope[heartdiseases$st_slope == "0"] = "upsloping"
heartdiseases$st_slope[heartdiseases$st_slope == "1"] = "flat"
heartdiseases$st_slope[heartdiseases$st_slope == "2"] = "downsloping"
heartdiseases$thalassemia[heartdiseases$thalassemia == "1"] = "normal"
heartdiseases$thalassemia[heartdiseases$thalassemia == "2"] = "fixed defect"
heartdiseases$thalassemia[heartdiseases$thalassemia == "3"] = "reversible defect"
heartdiseases$target[heartdiseases$target == "0"] = "no"
heartdiseases$target[heartdiseases$target == "1"] = "yes"
```

```{r}
str(heartdiseases)
```


```{r}
#Summarize the data
summary(heartdiseases)
```


#Simple Visualization for each Variable
Age Analysis 
```{r}
library(ggplot2)
library(ggpubr)
```

```{r}
ggplot(data = heartdiseases, aes(heartdiseases$age)) + 
  geom_histogram(aes(y = ..density..),
                 breaks=seq(20, 85, by = 2),
                 col = "black",
                 fill = "blue",
                 alpha=.2) +
  geom_density(col= "black") + 
  labs(title = "Histogram for Age", x = "Age", y = "Density")
```

```{r}
ggplot(heartdiseases, aes(age, fill = target)) +
        geom_density(alpha = 0.2) + 
        xlim (20, 85) +
        labs(title = "Heart Disease Frequency for Age", x = "Age", y = "Density")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
  
```

Most of the distribution of this dataset fall into between age 40 to age 70 (middle ages and elderly age range) 

Sex Analysis 
```{r}
hdsex = table(heartdiseases$sex, heartdiseases$target)
hdsex
```

```{r}
ggplot(heartdiseases, aes(sex)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for Sex", x = "Sex", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```
```{r}
table(heartdiseases$sex, heartdiseases$target)
```

Resting Blood Pressure
```{r}
ggplot(data = heartdiseases, aes(heartdiseases$resting_blood_pressure)) + 
  geom_histogram(aes(y = ..density..),
                 breaks=seq(80, 200, by = 2),
                 col = "black",
                 fill = "blue",
                 alpha=.2) +
  geom_density(col= "black") + 
  labs(title = "Histogram for Resting Blood Pressure", x = "Resting Blood Pressure", y = "Density")
```


Chest Pain Type
```{r}
hdcp = table(heartdiseases$chest_pain_type, heartdiseases$target)
hdcp
```

```{r}
#visualize the table
ggplot(heartdiseases, aes(chest_pain_type)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for Chest Pain Type", x = "Type", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

Cholesterol Analysis 
```{r}
ggplot(heartdiseases, aes(cholesterol, fill = target)) +
        geom_density(alpha = 0.2) + 
        xlim (100, 600) +
        labs(title = "Heart Disease Frequency for Cholesterol", x = "Cholesterol", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

Fasting Blood Sugar Anslysis 
```{r}
#visualize the table
ggplot(heartdiseases, aes(fasting_blood_sugar)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for Fasting Blood Sugar", x = "Fasting Blood Sugar", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

Resting Electrocardiographic Measurement Analysis 
```{r}
ggplot(heartdiseases, aes(rest_ecg)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for Resting Electrocardiographic Measurement Analysis", x = "Resting Electrocardiographic Measurement Analysis ", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

Maximum Heart Rate Achieved Analysis 
```{r}
ggplot(data = heartdiseases, aes(heartdiseases$max_heart_rate_achieved)) + 
  geom_histogram(aes(y = ..density..),
                 breaks=seq(50, 220, by = 2),
                 col = "black",
                 fill = "blue",
                 alpha=.2) +
  geom_density(col= "black") + 
  labs(title = "Histogram for Maximum Heart Rate Achieved", x = "Maximum Heart Rate Achieved", y = "Count")
```

St_Depression 
```{r}
ggplot(heartdiseases, aes(st_depression)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for St_Depression ", x = "St_Depression", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

The Number of Major Vessels 
```{r}
ggplot(heartdiseases, aes(num_major_vessels)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for The Number of Major Vessels", x = "The Number of Major Vessels", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

Thalassemia 
```{r}
ggplot(heartdiseases, aes(thalassemia)) +
  geom_bar(aes(fill=target)) + 
  labs(title = "Heart Disease Frequency for Thalassemia", x = "Thalassemia", y = "Count")+
        scale_fill_discrete(labels = c("No Disease", "Heart Disease"))
```

#Correlation Test 
```{r}
#subset the numerical variables 
heartdiseases1 = subset(heartdiseases, select = c(age, resting_blood_pressure, cholesterol, max_heart_rate_achieved, st_depression, num_major_vessels))
head(heartdiseases1)
```

```{r}
#create heat map to narrow down the significant correlation between variables.
hdcor <- round(cor(heartdiseases1),2)
hdcor
```

```{r}
#Melt the correlation matrix
library(reshape2)
melted_hdcor <- melt(hdcor)
head(melted_hdcor)
```

```{r}
#heatmap
library(ggplot2)
get_upper_tri <- function(hdcor){
  hdcor[lower.tri(hdcor)] <- NA
  return(hdcor)
}

upper_tri <- get_upper_tri(hdcor)
upper_tri
```


```{r}                                                                                          
melted_hdcor <- melt(upper_tri, na.rm = TRUE)

hdheatmap <- ggplot(data =  melted_hdcor, aes (Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1,
        size = 12, hjust = 1)) +
  coord_fixed()

hdheatmap + 
  geom_text (aes(Var2, Var1, label = value), color ="Black", size = 4) +
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
  
```
#Chi-Square Test for categorical values 
A chi-square test is used when you want to see if there is a relationship between two categorical variables. 

H0: There is no statistically significant relationship between independent variables and heart disease. 
H1: There is statistically significant relationship between independent variables and heart disease. 

We should reject H0 when P-value less than significance level 0.05. 

```{r}
#Sex
hdsex = table(heartdiseases$target, heartdiseases$sex)
chisq.test(hdsex)
#There is statistically significant relationship between sex and heart disease.
```

```{r}
#Chest Pain Type 
hdchest_pain_type = table(heartdiseases$target, heartdiseases$chest_pain_type)
chisq.test(hdchest_pain_type)
#There is statistically significant relationship between chest pain type and heart disease.
```

```{r}
#Fasting Blood Sugar
hdfasting_blood_sugar = table(heartdiseases$target, heartdiseases$fasting_blood_sugar)
chisq.test(hdfasting_blood_sugar)
#There is no statistically significant relationship between fasting blood sugar and heart disease.
```

```{r}
#Resting Electrocardiographic Measurement
hdrest_ecg = table(heartdiseases$target, heartdiseases$rest_ecg)
chisq.test(hdrest_ecg)
#There is statistically significant relationship between Resting Electrocardiographic Measurement and heart disease.
```

```{r}
#Excises induced angina
hdexcises_induced_angina = table(heartdiseases$target, heartdiseases$exercise_induced_angina)
chisq.test(hdexcises_induced_angina)
#There is statistically significant relationship between Excises induced angina and heart disease.
```

```{r}
#Thalassemia
hdthalassemia = table(heartdiseases$target, heartdiseases$thalassemia)
chisq.test(hdthalassemia)
#There is statistically significant relationship between fasting blood sugar and heart disease
```

#Logistic Regression Analysis 
```{r}
heartdiseases$target <- as.factor(heartdiseases$target)
str(heartdiseases)
```

```{r}
#split the data to train set and test set 
library(caTools)
split <- sample.split(heartdiseases, SplitRatio = 0.8)
split
train <- subset(heartdiseases, split == "TRUE")
test <- subset(heartdiseases, split == "FALSE")
```

```{r}
#Regression Model 1 
model1 <- glm (target ~ age + sex + chest_pain_type + resting_blood_pressure + cholesterol + fasting_blood_sugar + rest_ecg + max_heart_rate_achieved +  exercise_induced_angina + st_depression + num_major_vessels + thalassemia, data = train, family = 'binomial')
summary(model1)
```
Only sex, chest pain type (typical angina), resting blood pressure, max heart rate achieved, exercise induced angina, st_depression, and number of major vessels are statstically significant. Remove other variables. 


```{r}
#Regression Model 2 - update from model 1
model2 <- glm (target ~ sex + chest_pain_type + resting_blood_pressure + max_heart_rate_achieved + exercise_induced_angina + st_depression + num_major_vessels, data = train, family = 'binomial')
summary(model2)
```

Significant variables correlate to heart disease are: sex, chest pain type (typical angina),exercise induced angina, ST depression and max heart rate achieved and number of major vessels. 

```{r}
#Regression Model 3 - remove resting blood pressure 
model3 <- glm (target ~ sex + chest_pain_type + max_heart_rate_achieved + exercise_induced_angina + st_depression + num_major_vessels, data = train, family = 'binomial')
summary(model3)

```

#Regression diagnostic plots
```{r}
par(mfrow=c(2,2))
plot (model3)
```

#ANOVA Anlaysis 
H0: All variables have the same correlation to heart disease. 
H1: At least one variable do not have the same correlation to heart disease. 

Assume significant level at 0.05. 

Reject H0 when p-value is less than 0.05. 

```{r}
anova(model3, test = "Chisq")
```
All p-value are less than significant lever 0.05, we should reject H0. 
At least one variables do not have the same correlation to heart disease. 

```{r}
#model only the intercept
model3_1 <- glm(target ~ 1, data = train, family = "binomial")
#model with intercept + sex 
model3_2 <- glm(target ~ sex, data = train, family = "binomial")
#model with intercept + sex + chest pain type
model3_3 <- glm(target ~ sex + chest_pain_type, data = train, family = "binomial")
#model with intercept  + chest pain type + max_heart_rate_achieved
model3_4 <- glm(target ~ sex + chest_pain_type + max_heart_rate_achieved, data = train, family = "binomial")
#model with intercept  + chest pain type + resting blood pressure + max_heart_rate_achieved + exercise_induced_angina
model3_5 <- glm(target ~ sex + chest_pain_type + max_heart_rate_achieved + exercise_induced_angina, data = train, family = "binomial")
#model with intercept  + chest pain type + resting blood pressure + max_heart_rate_achieved + exercise_induced_angina + st_depression
model3_6 <- glm(target ~ sex + chest_pain_type + max_heart_rate_achieved + exercise_induced_angina + st_depression, data = train, family = "binomial")
#model containing all variabels 
model3_7 <- glm(target ~ sex + chest_pain_type + resting_blood_pressure + max_heart_rate_achieved  + st_depression + num_major_vessels , data = train, family = "binomial")
```

#ANOVA test between models 
Given the table above the model containing "site" doesn't provide any significant improvement (compared to NULL model) at the confidence level you mentioned

H0: Two models have the same significant effect while evaluating heart disease.
H1: Model 2 is significantly better than Model 1 while evaluating heart disease. 

Significant level at 0.05. Reject H0 when P-value is less than 0.05.

```{r}
anova(model3_1, model3_2, test = "LRT")
```

```{r}
anova(model3_2, model3_3, test = "LRT")
```

```{r}
anova(model3_3, model3_4, test = "LRT")
```

```{r}
anova(model3_4, model3_5, test = "LRT")
```

```{r}
anova(model3_5, model3_6, test = "LRT")
```

```{r}
anova(model3_6, model3_7, test = "LRT")
```

